Safe Memory Regions for Big Data Processing

### 2016-10-19 paper/type-inference.tex
moved discussion to appendix

\paragraph{Modularity Aspects.}
The type inference algorithm, as presented, traverses the entire program to
generate the set of constraints, which are solved en masse, using an iterative
fixed point computation. However, the type inference can be realized in a
modular and compositional fashion, subject only to the restrictions imposed
by recursion.

In the elaboration phase, we can process a class \C{C} only after any class
\C{B} that \C{C} depends on has been processed: class \C{C} depends on
class \C{B} if \C{B} is either \C{C}'s base class or the type of any field
of \C{C} depends on \C{B}. In effect, this means that any collection of
mutually recursive classes must be processed together. Non-recursive
dependences can be handled in a compositional fashion: if class \C{C}
depends on \C{B} non-recursively, then the elaboration can be done for
\C{B} first, and then \C{C} can be processed.

The same idea applies to the constraint-solving phase as well.
Given a set of constraints, we say that a predicate variable $\varphi_1$
\emph{directly-depends} on another predicate variable $\varphi_2$ if the set of
constraints includes a constraint $\isvalid{\varphi_1 \conj \phictxt}{F(\varphi_2)}$.
We say that $\varphi_1$ \emph{depends} on $\varphi_2$ if $\varphi_1$ transitively
depends on $\varphi_2$.
The constraint solver needs to process any collection of mutually dependent
predicate variables together.
In effect, this requires the type inference to process any collection of
mutually recursive methods together.
However, methods that are not mutually recursive can be processed separately.

@@

We now state the soundness theorem, which follows from
Theorems~\ref{thm:constraint-generation-sc} and~\ref{thm:constraint-solver-sc}
as well as the following observation about the elaboration phase.
For any program $p \in \absof{\FB}$, the elaboration phase produces a $q$
such that for any assignment $\soln$, we have $\absof{q[\soln]} = p$.

\begin{theorem}
\emph{(\textbf{Soundness})}
For any $p \in \absof{\FB}$, if type inference succeeds, it returns a $t$ such that
% Infer($p$) returns Some($t$), then
(1) $\absof{t} = p$, and
(2) $t$ is well-typed.
\end{theorem}

@@

However, in this context, completeness is more interesting than soundness.
(Note that an unsound inference algorithm can be made sound by simply type-checking
the resulting program.)
Theorems~\ref{thm:constraint-generation-sc} and~\ref{thm:constraint-solver-sc}
show that the second and third steps of the type-inference algorithm are complete.
%% We believe that the converse of the above lemma can also be established:
%% any substitution for the region variables and the predicate variables that makes the
%% expression well-typed will also satisfy the generated constraints.
%% Thus, if there is any possible instantiation (of the region variables and predicate variables) of
%% the elaborated program produced by the first phase that type checks, then the generated
%% constraints are satisfiable, and the second phase will identify it.
%% 
The only sources of incompleteness in the type inference algorithm are some heuristic
choices made in the first step (the elaboration phase), as explained below.

>>>>

@@

Theorems~\ref{thm:constraint-generation-sc} and~\ref{thm:constraint-solver-sc}
show that the second and third steps of the type-inference algorithm are both
sound and complete.
It is easy to verify the soundness of the first step (the elaboration phase):
For any program $p \in \absof{\FB}$, the elaboration phase produces a $q$
such that for any assignment $\soln$, we have $\absof{q[\soln]} = p$.
%
The soundness of the type inference algorithm follows.
The only sources of incompleteness in the type inference algorithm are some heuristic
choices made in the first step, as explained below.

%% \begin{theorem}
%% \emph{(\textbf{Soundness})}
%% For any $p \in \absof{\FB}$, if type inference succeeds, it returns a $t$ such that
%% % Infer($p$) returns Some($t$), then
%% (1) $\absof{t} = p$, and
%% (2) $t$ is well-typed.
%% \end{theorem}

@@

%% However, in this context, completeness is more interesting than soundness.
%% (Note that an unsound inference algorithm can be made sound by simply type-checking
%% the resulting program.)
%% Theorems~\ref{thm:constraint-generation-sc} and~\ref{thm:constraint-solver-sc}
%% show that the second and third steps of the type-inference algorithm are complete.
